Testing Logistic Regression Models (Supplement to: Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters)

### 2020-09-09 Scripts/Supplement_ModelTesting.Rmd
adding model testing data and changes to make Rmd more easily accessible without rerunning models denovo

#if loading data from previous model runs
load(file = "ModelTest_fits.Rdata")

@@

```{r, echo = F, message = F, warning = F}

@@

      # t_prior <- student_t(df = 7, location = 0, scale = 2.5)
      # #models fitting one intercept per unique area/taxon combination, plus
      # #a different slope for the effect of environmental variables on each taxon.

@@

      #compare model results
      loo_compare(waic(mod1.3vars), 
                  waic(mod1.2vars), 
                  waic(mod1.1vars.pH), 
                  waic(mod1.1vars.Temperature), 
                  waic(mod2.3vars), 
                  waic(mod2.2vars), 
                  waic(mod2.1vars.pH), 
                  waic(mod2.1vars.Temperature), 
                  waic(mod3.3vars), 
                  waic(mod3.2vars), 
                  waic(mod3.1vars.pH), 
                  waic(mod3.1vars.Temperature)
                  )

@@

```{r, echo = F, include = F}
#uncomment if running de novo; otherwise, load data from file, above
# stan.input <- data.for.stan %>% 
#       group_by(Area) %>% 
#       mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
#              pH = vegan::decostand(pH, method = "standardize"),
#              Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
#       group_by(taxa) %>%
#       mutate(Noccur = sum(binary)) %>%
#       filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
#       dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
#       ungroup() %>% 
#       unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
#       mutate(sampleIndex = 1:nrow(.),
#              taxonIndex = match(taxa, unique(taxa)),
#              areaIndex = match(Area, unique(Area)),
#              seasonIndex = match(Season, unique(Season)),
#              taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
#       dplyr::select(-AreaTaxa) 
# 
# finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
#                              data = stan.input, 
#                              family = "binomial",
#                              prior = t_prior,
#                              prior_intercept = t_prior, 
#                              chains = 4,
#                              adapt_delta = 0.99,
#                              refresh = 0)
```

@@

```{r, echo = F}
      stan.input %>% 
        filter(taxa == "Gonyaulacaceae|Alexandrium|NA") %>% 
        #filter(taxa %in% unique(taxa)[1:5]) %>% 
        dplyr::select(pH, Temperature, taxa, Area, binary, postpredict) %>% 
        pivot_longer(c(pH, Temperature)) %>% 
        ggplot(aes(x = value, y = binary)) +
        geom_point() +
        geom_point(aes(x = value, y = postpredict), color = "blue") +
        geom_smooth(aes(x = value, y = postpredict), color = "blue", method = "lm") +
        facet_grid(name ~ Area) +
        ylab("Prob of Presence") +
        xlab("pH (Standardized)") +
        ggtitle("Alexandrium")
```

>>>>

@@

```{r, echo = F, message = F, warning = F}
load("preliminaryModels.Rdata")

      t_prior <- student_t(df = 7, location = 0, scale = 2.5)
      #models fitting one intercept per unique area/taxon combination, plus
      #a different slope for the effect of environmental variables on each taxon.

@@

@@

      #compare model results
      loo_compare(waic(mod1.3vars), 
                  waic(mod1.2vars), 
                  waic(mod1.1vars.pH), 
                  waic(mod1.1vars.Temperature), 
                  waic(mod2.3vars), 
                  waic(mod2.2vars), 
                  waic(mod2.1vars.pH), 
                  waic(mod2.1vars.Temperature), 
                  waic(mod3.3vars), 
                  waic(mod3.2vars), 
                  waic(mod3.1vars.pH), 
                  waic(mod3.1vars.Temperature)
                  )
#optionally save model fits      
#save(list = ls(pattern = "mod"), file = "preliminaryModels.Rdata")      

@@

```{r, echo = F, include = F}
load("bestModelFit.Rdata") #if loading from existing data
stan.input <- data.for.stan %>%
      group_by(Area) %>%
      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
             pH = vegan::decostand(pH, method = "standardize"),
             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
      group_by(taxa) %>%
      mutate(Noccur = sum(binary)) %>%
      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>%
      ungroup() %>%
      unite(Area, taxa, col = "AreaTaxa", remove = F) %>%
      mutate(sampleIndex = 1:nrow(.),
             taxonIndex = match(taxa, unique(taxa)),
             areaIndex = match(Area, unique(Area)),
             seasonIndex = match(Season, unique(Season)),
             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
      dplyr::select(-AreaTaxa)

#uncomment if running de novo; otherwise, load data from file, above
# finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex),
#                              data = stan.input,
#                              family = "binomial",
#                              prior = t_prior,
#                              prior_intercept = t_prior,
#                              chains = 4,
#                              adapt_delta = 0.99,
#                              refresh = 0)
#optionally save
#save(finalModel, file = "bestModelFit.Rdata")
```

@@

```{r, echo = F, message = F, warning = F}
      stan.input %>% 
        filter(taxa == "Gonyaulacaceae|Alexandrium|NA") %>% 
        #filter(taxa %in% unique(taxa)[1:5]) %>% 
        dplyr::select(pH, Temperature, taxa, Area, binary, postpredict) %>% 
        pivot_longer(c(pH, Temperature)) %>% 
        ggplot(aes(x = value, y = binary)) +
        geom_point() +
        geom_point(aes(x = value, y = postpredict), color = "blue") +
        geom_smooth(aes(x = value, y = postpredict), color = "blue", method = "lm") +
        facet_grid(name ~ Area) +
        ylab("Prob of Presence") +
        xlab("pH (Standardized)") +
        ggtitle("Alexandrium")
```
